### 2020-12-21 ch01_introduction/introduction.tex
Added section on emergent constraints
Added introduction of chapter on ESMValTool

\section{Motivation}
\label{ch:01:motivation}

@@

- natural vs. anthropogenic effects (D \& A) -> clear evidence that climate
change over 21st century is caused by human influence
(https://globalwarmingindex.org/, Haustein et al., 2017)

\section{Structure of the thesis}
\label{sec:01:structure}

Parts of this thesis are published in multiple peer-reviewed publications (two
first-author studies and six co-author studies). If applicable, this is clearly
stated at the beginning of each chapter. \Cref{ch:02:scientific_background}
introduces the scientific background for this thesis. This includes relevant
literature that is used as a baseline for this thesis. \Cref{ch:03:esmvaltool}
gives an overview over the contributions made to the \ac{ESMValTool}, an
open-source software for the analysis of \acp{ESM}. These contributions helped
improving the routine evaluation of \acp{ESM} which is useful for the whole
scientific community and lead to co-authorship in four peer-reviewed studies
\autocite{Righi2020, Eyring2020, Lauer2020, Weigel2020}.
\Cref{ch:04:assessment_climate_metrics} covers the assessment of
policy-relevant climate metrics like the \ac{ECS} and the \ac{TCR} in the
latest generation of \acp{ESM}. This work is already published in two
scientific publications \autocite{Bock2020, Meehl2020}. Since the \ac{ECS} and
\ac{TCR} are considerably higher in this new climate model generation,
\cref{ch:05:paper_ecs} describes the assessment of emergent constraints (a
technique to reduce uncertainties in climate model projections, see
\vref{sec:02:techniques}) on the \ac{ECS} for these \acp{ESM}. The contents of
this chapter are published in \emph{Earth System Dynamics}
\autocite{Schlund2020a}. \Cref{ch:06:paper_gpp} focuses on a new method to
reduce climate model uncertainties based on \ac{ML}. As an example, the method
is applied to the photosynthesis rate at the end of the \nth{21} century, which
is already published in the \emph{Journal of Geophysical Research:
  Biogeosciences} \autocite{Schlund2020}. Finally, \cref{ch:07:summary_outlook}
provides a summary of the results of this thesis and gives an outlook of
possible future works.

>>>>

\section{Motivation}
\label{sec:01:motivation}

@@

- natural vs. anthropogenic effects (D\&A) -> clear evidence that climate
change over 21st century is caused by human influence
(https://globalwarmingindex.org/, Haustein et al., 2017)

- remove description of emergent constraints from section 1.3, since this will
probably be included in this section.

- answer question: what is climate sensitivity? What is \ac{ECS}? What is
\ac{TCR}? What is \ac{CMIP}?

- One sentence on \ac{ML}.

\section{Research Questions}
\label{sec:01:research_questions}

This section will present the \textbf{AWESOME} research questions!!!

\begin{enumerate}
  \item Did climate sensitivity change in the latest generation of climate
  models from \acs{CMIP}6 and can we see progress compared to older climate
  model generations?

  \item Can we use emergent constraints established in \acs{CMIP}3 and
  \acs{CMIP}5 to reduce uncertainties in climate sensitivity in the \acs{CMIP}6
  ensemble?

  \item Can we use \ac{ML}-based approaches to constrain gridded projections of
  the future climate with observations?
\end{enumerate}

\section{Structure of the Thesis}
\label{sec:01:structure}

Parts of this thesis are already published in multiple peer-reviewed
publications (two first-author studies and six co-author studies). Unless
stated otherwise, results, discussions and conclusions presented from these
studies originate from the author of this thesis. A detailed list of
corresponding references is given at the beginning of the relevant chapters.

This thesis is structured as follows: \Cref{ch:02:scientific_background}
introduces the scientific background. This includes relevant literature that is
used as a baseline for this thesis. \Cref{ch:03:esmvaltool} gives an overview
over the contributions made to the \ac{ESMValTool}, an open-source software for
the analysis of \acp{ESM}. These contributions helped improving the routine
evaluation of \acp{ESM} which is useful for the whole scientific community and
lead to co-authorship in four peer-reviewed studies \autocite{Eyring2020,
  Lauer2020, Righi2020, Weigel2020}. \Cref{ch:04:assessment_climate_metrics}
covers the assessment of policy-relevant climate metrics like the \ac{ECS} and
the \ac{TCR} in the latest generation of \acp{ESM}. This work is already
published in two scientific publications \autocite{Bock2020, Meehl2020}. Since
the \ac{ECS} and \ac{TCR} are considerably higher in this new climate model
generation, \cref{ch:05:paper_ecs} describes the assessment of emergent
constraints (a technique to reduce uncertainties in multi-model climate
projections, see \cref{subsec:02:emergent_constraints}) on the \ac{ECS} for
these \acp{ESM}. The contents of this chapter are published in \emph{Earth
  System Dynamics} \autocite{Schlund2020a}. \Cref{ch:06:paper_gpp} focuses on a
new method to reduce uncertainties in gridded multi-model projections of the
future climate with observations based on \ac{ML}. As an example, the method is
applied to the photosynthesis rate at the end of the \nth{21} century, which is
already published in the \emph{Journal of Geophysical Research: Biogeosciences}
\autocite{Schlund2020}. Finally, \cref{ch:07:summary_outlook} provides a
summary of the results of this thesis and gives an outlook of possible future
works.

### 2020-12-21 ch02_scientific_background/scientific_background.tex
Added introduction of chapter on ESMValTool

This chapter introduces the scientific background of this thesis. First, basic
concepts of climate model simulations and associated uncertainties are
introduced. Next, the fundamental biogeochemical processes of the global carbon
cycle and important metrics describing climate change are presented. Finally,
state-of-the art techniques used reduce uncertainties in projections of the
future climate are shown. These methods form the basis for the new techniques
developed in this thesis.

>>>>

This chapter introduces the scientific background of this thesis. First, basic
concepts of climate model simulations and associated uncertainties are
introduced. Next, important metrics describing climate sensitivity and
fundamental biogeochemical processes of the global carbon cycle are presented.
Finally, state-of-the art techniques that can be used to reduce uncertainties
in projections of the future climate with observations are shown. These methods
form the basis for new techniques developed in this thesis.

### 2020-12-21 ch03_esmvaltool/esmvaltool.tex
Initialized chapters
Added first figure
Added further text and figures to ch.2
Added section 2.2
Started section on techniques for reducing uncertainties in climate model projections
Added introduction of chapter on ESMValTool

%=============================================================================%
%                                Dissertation                                 %
%                               Manuel Schlund                                %
%                                  (c) 2020                                   %
%=============================================================================%
%                            Scientific Background                            %
%=============================================================================%

\chapter{Scientific Background}
\label{ch:scientific_background}

TBA.

>>>>

%=============================================================================%
%                                Dissertation                                 %
%                               Manuel Schlund                                %
%                                  (c) 2020                                   %
%=============================================================================%
%                                 ESMValTool                                  %
%=============================================================================%

\chapter{Improving Routine Climate Model Evaluation}
\label{ch:03:esmvaltool}

In order to answer the research questions of this thesis posed in
\cref{sec:01:research_questions}, a reliable tool to read, process and evaluate
output from climate model simulations is necessary. A valuable software that
fits these criteria is the \ac{ESMValTool}. The \ac{ESMValTool} is an
open-source community diagnostics and performance metrics tool for the routine
evaluation of \ac{ESM} output, which particularly facilitates the analysis of
\ac{CMIP} models (\url{https://www.esmvaltool.org}). For this reason, all
analyses presented in
\cref{ch:04:assessment_climate_metrics,ch:05:paper_ecs,ch:06:paper_gpp} of this
thesis were implemented into the \ac{ESMValTool}. Apart from that, further
substantial contributions to the code base of the \ac{ESMValTool} have been
made, which lead to co-authorship in the four peer-reviewed studies that
provide its scientific documentation \autocite{Eyring2020, Lauer2020,
  Righi2020, Weigel2020}. After a brief overview over the \ac{ESMValTool}'s
structure, this chapter presents all these contributions that are not
documented in other chapters of this thesis.

\section{The \acf{ESMValTool}}
\label{sec:03:esmvaltool}

- overview ESMValTool

- CMOR

- Changes in ESMValCore + table

- Added recipes and diagnostics for ESMValTool + table

* Anav13

* MDER

* TCR

* ECS

* Schlund20jgr

* Schlund20ESD

* Flato13

- total number of changed lines of code
